# Entering data

Good data organization is the foundation of any research project, 
and this begins with entering and archiving data. 
There are several reasonable options for data entry, for example:

1. Spreadsheet
2. Text file
3. Database
4. Form (web or GUI databases). 

In R, you can also input data directly, which can be the right choice if you are working with a small dataset. 
Here, you will learn how to enter the data from your experiments
and save them using R. 

Let's start by
making table 1 for your report, which we will call 
`solution-concentration-effect-on-potato.csv`:

```{r}
table1 <- data.frame(weight = c("initial", "final", "difference", "percent-change"), 
                     NaCl-0-percent = c(, , ,), 
                     NaCl-0.9-percent = c(, , ,), 
                     NaCl-2-percent = c(, , ,), 
                     NaCl-5-percent = c(, , ,), 
write.csv(x = table1, file = "solution-concentration-effect-on-potato.csv", row.names = FALSE)
```

Notice that we used the file extension ".csv".  
CSV stands for comma-separated values, which is a format for 
tabular data stored in a text file where a comma separates the columns of the data. 
That is why a comma separates each of the values in the rows for the different solutions. 

Keep in mind that the values entered into the tables should be 
decimals greater than or equal to zero, and likely less than
100 since you measured mass in grams and the potato
cubes for the experiments were 8cm^3 in size. 

Now, we will make table 2, which includes the class averages and we will call 
`solution-concentration-effect-on-potato-averages.csv`:

```{r}
table2 <- data.frame(weight = c("percent-change", "minimum-value", "maximum-value", "difference"), 
                     NaCl-0-percent = c(, , ,), 
                     NaCl-0.9-percent = c(, , ,), 
                     NaCl-2-percent = c(, , ,), 
                     NaCl-5-percent = c(, , ,), 
write.csv(x = table1, file = "solution-concentration-effect-on-potato-averages.csv", row.names = FALSE)
```

We can load our tables into R the following way:

```{r}
table1 <- read.csv(file = "solution-concentration-effect-on-potato.csv")
table2 <- read.csv(file = "solution-concentration-effect-on-potato-averages.csv")
table1
table2
```

Here, we used the function `read.csv` for reading in our data files
in CSV format.
